An Empirical Analysis of C Preprocessor Use

### 2000-02-20 macros/papers/emp-use-2.tex
Changes for TSE final copy (which has now been sent off).

% $Id$
\documentclass[10pt]{article}
\def\typeofdocument{article}    % ``article'' or ``report''
\usepackage{epsfig}
\usepackage{dcolumn}
\usepackage{alltt}
\usepackage{rotating}

@@

\date{% Revision of Technical Report UW-CSE-97-04-06 (original date April 22, 1997) \\
Department of Computer Science and Engineering \\
University of Washington \\
Box 352350, Seattle, WA  98195-2350  USA \\
{\small \{{\tt mernst},{\tt gjb},{\tt notkin}\}{\tt @cs.washington.edu}} \\
18 January 2000}

@@

%\noindent
%Keywords: C preprocessor, Cpp, C, C++, program understanding, file
%inclusion, macro, macro substitution, conditional compilation, empirical study

@@

Section~\ref{sec:background} provides additional detail about the
difficulties imposed by the preprocessor.
Section~\ref{sec:methodology} describes our experimental methodology.
Sections~\ref{sec:first-content-section}--\ref{sec:last-content-section}
present the bulk of results about macro preprocessor use.
Section~\ref{sec:related} discusses related work, and
Section~\ref{sec:easier-to-understand} suggests techniques for mitigating
the negative impact of Cpp on program understanding.
Section~\ref{sec:future-work} presents avenues for future work, and the
concluding section discusses the relevance of the research.

@@

\begin{figure}
% This works, but the figures are upside-down.
% \centerline{\epsfig{file=fig/exp-dep-bymacro.eps,angle=90,height=3.75in}}
% \centerline{\epsfig{file=fig/incl-dep-bymacro.eps,angle=90,height=3.75in}}
% Cannot use ``height'' when rotating by -90 or +270; I do not know why.
% \centerline{\epsfig{file=fig/exp-dep-bymacro.eps,angle=270,width=.5\linewidth}}
% \bigskip
% \centerline{\epsfig{file=fig/incl-dep-bymacro.eps,angle=270,width=.5\linewidth}}
% When Mike exports Excel charts, angle is 270.  When Greg exports, angle is 0.
\centerline{\epsfig{file=fig/exp-dep-bymacro.eps,angle=270,width=.48\linewidth}%
\ %
\epsfig{file=fig/incl-dep-bymacro.eps,angle=270,width=.48\linewidth}}
%%Numbers: read off of Excel exp-dep-bymacro
\captionsmall{Dependences by macro name for \numdependmacronames\ macro
  names in \numdependpackages\ packages.  The first bar represents all
  macros that are completely unused in a package.
  Each other bar represents all macros
  that control a specified percentage range of the lines in a
  package.  For instance, the 0.17--0.26 bar in the
  expansion dependence chart indicates that 257 macros each control between
  0.17\% and 0.26\% of the entire package that contains that macro.  The
  maximum falls in the penultimate bucket (i.e., the rightmost bucket is
  the first empty one).  The second bar of each graph represents all macros
  controlling at least one line ($\varepsilon$ is a very small but nonzero
  value), but no more than .01% of the entire package.  A log scale is used
  for the x axis.}

@@

% Not really right:  Do not want the ``References'' section head to be small.
{\small \bibliography{evil}}

>>>>

% $Id$
% Process with LaTeX2e
\documentclass[10pt]{article}
\def\typeofdocument{article}    % ``article'' or ``report''
\usepackage{epsfig}
\usepackage{dcolumn}
\usepackage{alltt}
\usepackage{rotating}

@@

\date{% Revision of Technical Report UW-CSE-97-04-06 (original date April 22, 1997) \\
Department of Computer Science and Engineering \\
University of Washington \\
Box 352350, Seattle, WA  98195-2350  USA \\
{\small \{{\tt mernst},{\tt gjb},{\tt notkin}\}{\tt @cs.washington.edu}}
%  \\ 18 January 2000
}

@@

\noindent
Keywords: C preprocessor, Cpp, C, C++, macro, macro substitution, file
inclusion, conditional compilation, empirical study, program understanding

@@

The remainder of this \typeofdocument{} is organized as follows.
Section~\ref{sec:background} provides additional detail about the
difficulties imposed by the preprocessor.
Section~\ref{sec:methodology} describes our experimental methodology.
Sections~\ref{sec:first-content-section}--\ref{sec:last-content-section}
present the bulk of results about macro preprocessor use.
Section~\ref{sec:related} discusses related work, and
Section~\ref{sec:easier-to-understand} suggests techniques for mitigating
the negative impact of Cpp on program understanding.
Section~\ref{sec:future-work} presents avenues for future work, and the
concluding section discusses the relevance of the research.

@@

\begin{figure}
% This works, but the figures are upside-down.
% \centerline{\epsfig{file=fig/exp-dep-bymacro.eps,angle=90,height=3.75in}}
% \centerline{\epsfig{file=fig/incl-dep-bymacro.eps,angle=90,height=3.75in}}
% Cannot use ``height'' when rotating by -90 or +270; I do not know why.
% \centerline{\epsfig{file=fig/exp-dep-bymacro.eps,angle=270,width=.5\linewidth}}
% \bigskip
% \centerline{\epsfig{file=fig/incl-dep-bymacro.eps,angle=270,width=.5\linewidth}}
% When Mike exports Excel charts, angle is 270.  When Greg exports, angle is 0.
\centerline{\epsfig{file=fig/exp-dep-bymacro.eps,angle=270,width=.48\linewidth}%
\ %
\epsfig{file=fig/incl-dep-bymacro.eps,angle=270,width=.48\linewidth}}
%%Numbers: read off of Excel exp-dep-bymacro
\captionsmall{Dependences by macro name for \numdependmacronames\ macro
  names in \numdependpackages\ packages.  The first bar represents all
  macros that are completely unused in a package.
  Each other bar represents all macros
  that control a specified percentage range of the lines in a
  package.  For instance, the 0.17--0.26 bar in the
  expansion dependence chart indicates that 257 macros each control between
  0.17\% and 0.26\% of the entire package that contains that macro.  The
  maximum falls in the penultimate bucket (i.e., the rightmost bucket is
  the first empty one).  The second bar of each graph represents all macros
  controlling at least one line ($\varepsilon$ is a very small but nonzero
  value), but no more than .01\% of the entire package.  A log scale is used
  for the x axis.}

@@

\bibliography{evil}

\noindent
\epsfig{file=fig/ernst-photo.eps,width=1.4in}

\textbf{Michael D. Ernst} will receive the Ph.D. in Computer Science and
Engineering from the University of Washington in June 2000.  He has
previously been a lecturer at Rice University and a researcher at Microsoft
Research.  He holds the S.B. and S.M. degrees from MIT.  Ernst's primary
technical interest is programmer productivity, encompassing software
engineering, program analysis, compilation, and programming language
design.  However, he has also published in artificial intelligence, theory,
and other areas of computer science.

\bigskip
\noindent
\epsfig{file=fig/badros-photo.eps,width=1.4in}

\textbf{Greg J. Badros} is a final-year Ph.D. candidate at the University of
Washington in Seattle, USA where he earned his M.Sc.\ degree in 1998.  He
graduated \textit{magna cum laude} with a B.S.  degree in Mathematics and
Computer Science from Duke University in 1995.  He is the primary author of
the Scheme Constraints Window Manager and the Cassowary Constraint Solving
Toolkit.  His research interests include constraint technology, software
engineering, languages, and the Internet.

\bigskip
\noindent
\epsfig{file=fig/notkin-photo.eps,width=1.4in}

\textbf{David Notkin} 
is the Boeing Professor of Computer Science and Engineering at
the University of Washington. Before joining the faculty in 1984, he
received his PhD degree at Carnegie Mellon University in 1984 and his ScB
degree at Brown University in 1977.  Dr.~Notkin received the National Science
Foundation Presidential Young Investigator Award in 1988; served as the
program chair of the First ACM SIGSOFT Symposium on the Foundations of
Software Engineering; served as program co-chair of the 17th International
Conference on Software Engineering; chaired the Steering committee of the
International Conference on Software Engineering (1994-1996); served as
charter associate editor of both ACM Transactions on Software Engineering
and Methodology and the Journal of Programming Languages; serves as an
Associate Editor of the IEEE Transactions on Software Engineering; was named
as an ACM Fellow in 1998; and serves as the chair of ACM SIGSOFT\@.  His
research interests are in software engineering in general and in software
evolution in particular.  Dr.~Notkin is a senior member of the IEEE.
